Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-11-26 inst/paper/main.tex
intro

To further illustrate the challenges at hand, consider the interaction
between a static analysis tool and a dynamic language. An abstract
interpretation-based program analyzer computes an over-approximation
of the set of possible behaviors exhibited by the system under
study~\cite{cc77}. The presence of a reflective operation entails that
the operation may have any behavior that can be expressed in the
language; as if we replaced it by any legal sequence of instructions.
As dynamic languages tend to be permissive, the analysis has to assume
that many (and sometimes all) functions in scope may have been
redefined, e.g. that \texttt{`+`} now opens an Internet connection or
something equally surprising. This means that a single occurrence of
\eval will cause the static analyzer to loose all information about
the program state and the meaning of identifiers. This loss of
precision can sometimes be mitigated by analyzing the string passed to
\eval~\cite{} to bound its possible behavior but when the string comes
from outside the program not much can be done.  A group researchers
argued giving up on soundness and, instead, under-approximate dynamic
features~\cite{soundy} In their words ``a practical analysis,
therefore, may pretend that \eval does nothing, unless it can
precisely resolve its string argument at compile time.''  Alas
assuming that \eval does not have side-effects may be unduly
optimistic.

While there has been previous work investigating the usage of \eval in web
programming and JavaScript~\cite{ecoop11}, there are no other studies that
shed light on the usage of \eval in other programming languages and
application domains.  It is thus reasonable to wonder if the results from
previous work can be carried over to other contexts. For this paper, we have
picked the data science language R~\cite{r} as our object of study. R is
both different from JavaScript in linguistic terms and in terms of its user
base. JavaScript was designed to run untrusted code in browser, while R was
designed for statistical computing on a desktop. JavaScript is used as a
general purpose language by a wide community of programmers; while R is used
for scientific computing by data scientists and domain experts with, often,
limited programming experience.

>>>>

To further illustrate the challenges at hand, consider the interaction
between a static analysis tool and a dynamic language. An abstract
interpretation-based program analyzer computes an over-approximation
of the set of possible behaviors exhibited by the system under
study~\cite{cc77}. The presence of a reflective operation entails that
the operation may have \emph{any} behavior that can be expressed in
the target language; i.e. as if we replaced it by any legal sequence
of instructions.  As dynamic languages tend to be permissive, the
analysis has to, for example, assume that many (and sometimes all)
functions in scope may have been redefined, e.g. that \texttt{`+`} now
opens an Internet connection or something equally surprising. This
means that a single occurrence of \eval causes the static analyzer to
loose all information about the program state and the meaning of
identifiers. This loss of precision can sometimes be mitigated by
analyzing the string passed to \eval~\cite{} to bound its possible
behavior but when the string comes from outside the program not much
can be done.  A frustrated group researchers argued giving up on
soundness and, instead, under-approximating dynamic
features (soundiness)~\cite{soundy}. In their words ``a practical analysis,
therefore, may pretend that \eval does nothing, unless it can
precisely resolve its string argument at compile time.''  Alas,
assuming that \eval does not have side-effects, or that side-effects
will not affect the results of the analysis, may be unduly optimistic.

\vspace{2mm}\noindent\emph{Is Past Prologue?} Previous work
investigated how \eval is used in web programming, specifically in web
sites that use JavaScript~\cite{pldi10a}. In 2010, 17 of the largest
web site used the feature. In 2011, 82\% of the 10,000 most accessed
sites used \eval~\cite{ecoop11}.  Yet, the strings passed to \eval,
and their behaviors when executed, are far from random; it was shown
that when one can observe several calls to \eval, the ``shape'' of
future calls can be predicted with 97\% accuracy~\cite{oopsla12b}.
Overall, practical usage suggested that most reflective calls were
relatively harmless. While this backs up the soundiness squad's
approach, does it generalize to other application domains than web
programming and to other languages?

application domains.  It is thus reasonable to wonder if the results from
previous work can be carried over to other contexts. For this paper, we have
picked the data science language R~\cite{r} as our object of study. R is
both different from JavaScript in linguistic terms and in terms of its user
base. JavaScript was designed to run untrusted code in browser, while R was
designed for statistical computing on a desktop. JavaScript is used as a
general purpose language by a wide community of programmers; while R is used
for scientific computing by data scientists and domain experts with, often,
limited programming experience.
